---
title: "Fire data exploration"
author: "Sam Morris"
date: "10 June 2015"
output: pdf_document
---

# EDA for Georgia fire data

```{r setup, echo=FALSE, include=FALSE}
library(fields)
library(maps)
source(file = "../code/analysis/fire/adj.R", chdir = TRUE)
load(file = "../code/analysis/fire/gaCntyFires.RData")
load(file = "../code/analysis/fire/fire_data.RData")
load(file = "../code/analysis/fire/chi.RData")
```

```{r chiplot-setup, echo = FALSE}
# sort the adjacency matrix
order <- rep(NA, ncol(Y))
order[1] <- 1
for (i in 1:length(order)) {
  this <- order[i]
  neighbors <- which(ADJ[this, ] == 1)
  for (k in 1:length(neighbors)) {
    if (!(neighbors[k] %in% unique(order))) {
      this <- min(which(is.na(order)))
      order[this] <- neighbors[k]
    }
  }
}

chi.sort <- matrix(NA, ncol(Y), ncol(Y))
for (i in 1:ncol(Y)) {
  for (j in i:ncol(Y)) {
    pair.i <- order[i]
    pair.j <- order[j]
    chi.sort[i, j] <- chi.sort[j, i] <- chi[pair.i, pair.j]
  }
}

chi.sort <- ifelse(chi.sort < 0, 0, chi.sort)
# image.plot(1:ncol(Y), 1:ncol(Y), chi, main = "estimated chi")
```

## Dependence

I used the `chiplot` function in `evd` to estimate the pairwise $\chi$ for all pairs of counties.
The value we're using is the average of $\chi(u)$ for $u = 0.909, 0.917, 0.925, 0.933, 0.940, 0.948$.
Then to get the estimate for $\theta$, I'm using $2 - \chi$.
Currently, I am estimating $\theta$ only using the annual forest fire burn amount, but I will be changing this to work with the daily time series.
This plot is an estimate of $\widehat{\theta}$ and not $\widetilde{\theta}$.
I plan to implement the method described in the draft manuscript as the next step to get $\widehat{\theta}$ and $\widetilde{\theta}$.

```{r chiplot, echo = FALSE, fig.height=3.5, fig.width=5}
image.plot(1:ncol(Y), 1:ncol(Y), chi.sort, main = "estimated chi")
```

```{r ecplot, echo = FALSE, fig.height=3.5, fig.width=5}
image.plot(1:ncol(Y), 1:ncol(Y), 2 - chi.sort, main = "estimated EC")
```

## How best to sort the sites?

I did use the adjacency matrix to try to rearrange the counties in the plots so neighboring counties are near each other, but this still needs to be refined some more.
I tried to figure out how to identify the latitude and longitude of the centroids of the counties, but I could not find anything.
Do you have any code that takes the information from the `maps` package and can calculate the centroids?
